---
title: "Avengers Endgame text wrangling"
author: "Lory Salazar"
date: "2/24/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE )
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(here)
```

## Introduction: Text wrangling and analysis for the screenplay of the movie Avengers: Endgame

In this report, we use text mining and analysis skills to look at the top used words in the screenplay for Avengers: Endgame. Additionally, we perform sentiment analysis using ``

```{r}
endgame_text <- pdf_text(here("endgame.pdf"))
```

```{r}
# Creating a tidy dataframe
endgame_tidy <- data.frame(endgame_text) %>% 
  mutate(text_full = str_split(endgame_text, pattern = "\\n")) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) %>% 
  slice(-(1:8))

endgame_tokens <- endgame_tidy %>% 
  unnest_tokens(word, text_full) %>% 
  select(-endgame_text)

# Remove all stopwords, then look at what the counts look like
endgame_nonstop <- endgame_tokens %>% 
  anti_join(stop_words)

nonstop_counts <- endgame_nonstop %>% 
  count(word)
```


```{r}
# Looking at the top most used words in the screenplay
top_5 <- nonstop_counts %>% 
  arrange(-n) %>% 
  slice(1:5)

# Because our word cloud will have the top 100 most used words, first create a subset of the top 100 words
top_100 <- nonstop_counts %>% 
  arrange(-n) %>% 
  slice(1:100)


```

### Word cloud of top most used words in Avengers: Endgame
```{r}
# Creating a word cloud of the top words used 

ggplot(data = top_100, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "diamond") +
  scale_size_area(max_size = 6) +
  scale_colour_gradientn(colors = c("cyan3","blue","purple")) +
  theme_minimal() +
  labs(title = "Top 100 words in Avengers: Endgame")

```

**Figure 1.** A wordcloud to visualize the top 100 words used in the screenplay for Avengers: Endgame. Bigger words indicate a more frequently used word. Purple, blue, and cyan colors indicate a word appears more to less frequently, respectively.

### Using sentiment analysis to visualize the sentiments of Avengers: Endgame
```{r}
# First looking at what the words look like in "afinn"
afinn_pos <-get_sentiments("afinn") %>% 
  filter(value > 2)

# Looking at "afinn" specifically four our text, then getting the counts
endgame_afinn <- endgame_nonstop %>% 
  inner_join(get_sentiments("afinn"))

afinn_counts <- endgame_afinn %>% 
  count(value)

# Performing the same steps, but for "nrc", which we will then use for our plot.

endgame_nrc <- endgame_nonstop %>% 
  inner_join(get_sentiments("nrc"))

nrc_counts <- endgame_nrc %>% 
  count(sentiment)

ggplot(data = nrc_counts, aes(x = n, y = sentiment)) +
  geom_col() +
  labs(title = "Sentiment analysis for Avengers: Endgame",
       x = "Number of occurences", y = "Sentiment") +
  theme_minimal()

```

**Figure 2.** The number of occurrences of words with different sentiments in Avengers: Endgame using the nrc lexicon.

### Data citation
Christopher Markus & Stephen McFeely
Markus, Christopher, and McFeely, Stephan. Screenplay of *Avengers: Endgame*. Disney Studio Awards, https://disneystudiosawards.s3.amazonaws.com/avengers-endgame/Avengers_Endgame.pdf